---
title: "Case study_sleep deprivation"
author: "Wenjing"
date: "2021/3/8"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Wenjing Wang r0818232 
# Statistical methods for bioinformatics_Random slope model

# 0. Setout
```{r echo = TRUE，eval = TRUE}
setwd("D:/Msc_Bioinformatics/1st_2s/Statistical Methods for Bioinformatics/Part1/Projects/Project6_Sleep deprivation/")
getwd

## Get the data
Rea_Object <- read.table("sleep.txt",header = T,sep = "")

## Attach data to the search path
attach(Rea_Object)
```


# 1.	Change of reaction time for the subject over the days of sleep deprivation: Spaghetti-plot

```{r echo = TRUE，eval = TRUE}
n = length(unique(Subject))
interaction.plot(Days,Subject,Reaction, xlab = "Days", ylab = "Reaction time",legend = F)
lm.reg <- lm(Reaction~Days)
lm.reg
abline(lm.reg,col = "red",lty = 3)
```


# 2. Data of means per subject: Data table

```{r echo = TRUE，eval = TRUE}
## Mean
Rea_Object.mean = tapply(Reaction, Subject, mean)
Rea_Object.mean

## Standard deviation
Rea_Object.sd = tapply(Reaction, Subject, sd)
Rea_Object.sd

## Variance
Rea_Object.var = tapply(Reaction, Subject, var)
Rea_Object.var

## Number:
Rea_Object.n = table(Subject)
Rea_Object.n
```

# 3. Changes of reaction time over subject and days: Box plots
```{r echo = TRUE，eval = TRUE}
par(mfrow = c(2,1)) # Separated picture
boxplot(Reaction~Subject,xlab = "Subject",ylab = "Reaction time") ##  Changes of reaction time over subject
boxplot(Reaction~Days,xlab = "Days",ylab = "Reaction time") ##  Changes of reaction time over Days
```

# 4. Linear regression per subject: Scatterplots

```{r echo = TRUE，eval = TRUE}
cf <- sapply(Rea_Object$Subject, function(x) coef(lm(Reaction~Days,data = subset(Rea_Object,Subject == x))))
##  Calculate the linear regression(Reaction~Days) per subject 

Sx <- reorder(Rea_Object$Subject,cf[1,]) ## Sorted by the subject index

library(lattice)
xyplot(Reaction~Days|Sx,data = Rea_Object, type = c('p','r'),auto.key = T, aspect = "xy", 
       par.settings = list(axis.text = list(cex = 0.6), fontsize = list(text = 8, points = 10)),
       scales = list(x = list(at=c(0,1,2,3,4,5,6,7,8,9),labels=c("0","1","2","3","4","5","6","7","8","9"))))
## Draw the scatterplots which show the linear regression plots for each subject
```

# 5. Coefficients and R square for subjects of reaction on days: Histograms

```{r echo = TRUE，eval = TRUE}
## Calculate coefficients(intercept & slope)
lin.reg.coef <- by(Rea_Object,Rea_Object$Subject,function(data) coef(lm(Reaction~Days,data = data)))
lin.reg.coef1 <- unlist(lin.reg.coef)
names(lin.reg.coef1) <- NULL
lin.reg.coef2 = matrix(lin.reg.coef1,length(lin.reg.coef1)/2,2,byrow = TRUE)

## Calculate the R squared
lin.reg.r.squared <- by(Rea_Object,Rea_Object$Subject,function(data) summary(lm(Reaction~Days,data = data))$r.squared)
lin.reg.r.squared1 <- as.vector(unlist(lin.reg.r.squared))

## Making Histograms 
par(mfrow = c(3,1))
hist(lin.reg.coef2[,1],xlab = "Intercept", col = "lightblue", main = "Histogram of Subject intercept")
hist(lin.reg.coef2[,2],xlab = "Slope", col = "lightblue", main = "Histogram of Subject Slope")
hist(lin.reg.r.squared1,xlab = "R squared", col = "lightblue", main = "Histogram of Subject R squared")
```

# 6.	Specify a random slope of Days for each subject: Data table

```{r echo = TRUE，eval = TRUE}
library(lme4)
library(lmerTest)

## Use the code (1+Days | Subject) to specify a random intercept for each subject, and a random slope of Days for each subject.
res <- lmer(Reaction ~ Days + (1 + Days|Subject), data = Rea_Object)

## Fixed effects section of the model summary
summary(res)

## The 95% coverage interval of the slope
confint(res,level = 0.95,method = "Wald")
```

# 7. Standardized residuals: Scatter plots

```{r echo = TRUE，eval = TRUE}
res.std <- resid(res)/sd(resid(res))
plot(res.std, ylab = "Standardized Residuals")
```


# 8. Detect the outlier: Diagnostic plots
```{r echo = TRUE，eval = TRUE}
car::outlierTest(res)
car::infIndexPlot(res)
```


# 9. The difference between the fitted and observed values: Lines and Sctterplots
```{r echo = TRUE，eval = TRUE}
library(ggplot2)
ggplot(as.data.frame(res.std),aes(sample = res.std))+geom_qq()+geom_qq_line()
```



